A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-27 arxiv/proof_main.tex
wip
sec 3 wip
sec 7.2

\section{Proof of Main Results}

\subsection{Two Tasks}\label{app_proof_main_thm}

%We shall refer to random matrices of the form $X^\top X$ as sample covariance matrices following the standard notations in high-dimensional statistics. The second lemma extends Lemma \ref{lem_minv} for a single sample covariance matrix to the sum of two independent sample covariance matrices. It is the main random matrix theoretical input of this paper.
%%which deals with the inverse of the sum of two random matrices, which
%%any is can be viewed as a special case of Theorem \ref{thm_model_shift}.
%
%\begin{lemma}[Variance bound: Lemma \ref{lem_cov_shift_informal} restated]\label{lem_cov_shift}
%	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
%	Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
%	Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and let $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M$ in descending order. Let $A$ be any $p\times p$ matrix that is independent of $X_1$ and $X_2$. We have that for any constant $\e>0$,
%	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
%	\begin{align}\label{lem_cov_shift_eq}
%		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}A} = \frac{1}{\rho_1+\rho_2}\frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} A} +\bigo{\|A\| p^{-1/2+\epsilon}}
%	\end{align}
%with high probability, where $(a_1, a_2)$ is the solution to the following deterministic equations:
%	\begin{align}
%		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
%	\end{align}
%\end{lemma}
%
%Finally, the last lemma describes the asymptotic limit of $(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}$, which will be needed when we estimate the first term on the right-hand side of \eqref{eq_te_mtl_2task}.

%We will give the proof of this lemma in Section \ref{sec_maintools}.

%\subsection{Proof for Two Tasks with General Covariates}\label{app_proof_main}

%To illustrate the idea, we observe that by using Lemma \ref{lem_minv}, we have that
%\[ \te(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{n_2 - p}\bigtr{\Sigma_2^{-1}}. \]
%We shall also derive the limit of $\te(\hat{\beta}_t^{\MTL})$.
%\todo{write a brief technical overview}

In this subsection, we give the proof Theorem \ref{thm_main_informal}. During the proof we shall also give explicit expressions for $\Delta_{\bias}$ and $\Delta_{\vari}$, which might be of interest to some readers. %We shall consider the case where the entries of $\e_1$ and $\e_2$ have the same variance $\sigma_1^2=\sigma_2^2=\sigma^2$. 
%Moreover, we shall denote $\beta_1$ and $\beta_2$ by $\beta_1$ and $\beta_2$, standing for ``source task" and ``target task", respectively. 
%First, we introduce several quantities that will be used in our statement, and they are also related to the quantities in Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}. Given the optimal ratio $\hat v$, 
%%let $\hat{M} = \hat{v} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix, and ${\hat\lambda}_1\ge {\hat\lambda}_2 \ge \dots \ge {\hat\lambda}_p$ be the eigenvalues of $\hat{M}^{\top}\hat{M}$. Define 
%let $(\hat a_1, \hat a_2)$ be the solution to the following system of deterministic equations,
%	\be
%		 \hat a_1 +  \hat a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad  \hat a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac1p\sum_{i=1}^p \frac{ \hat v^2 \lambda_i^2 \hat a_1}{ \hat v^2 \lambda_i^2\hat a_1 +  \hat a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.\label{eq_a2} \\
%		 \ee
%		 After obtaining $(\hat a_1,\hat a_2)$, we can solve the following linear equations to get $(\hat a_3,\hat a_4)$:
%\begin{gather}
%		\left(\rho_2 \hat a_2^{-2}- \hat b_0\right)\cdot \hat  a_3 - \hat b_1 \cdot \hat a_4
%		=\hat b_0, \quad \left(\rho_1\hat a_1^{-2} - \hat b_2  \right)\cdot \hat a_4 - \hat b_1 \cdot \hat a_3 =\hat b_1 .\label{eq_a3} 
%%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
%	\end{gather}
%where we denoted
%$$\hat b_k:= \frac1{p}\sum_{i=1}^p \frac{\hat v^{2k} \lambda_i^{2k}}{ (\hat a_2 +\hat v^2 \lambda_i^2\hat a_1)^2  },\quad k=0,1,2.$$
With $a_i$, $i=1,2,3,4$, given in Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}, 
we define the following matrix
\be\label{defnpihat}\Pi \equiv \Pi(\hat v)= \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot \hat v^2{M} \frac{(1 +   a_3)\id + \hat v^2 a_4 {M}^{\top} {M}}{( \hat v^2 a_1 {M}^{\top} {M}+ a_2 )^2} {M}^{\top}.\ee
%which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
%$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
Moreover, we introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=\left(1- \rho_1^{-1/2}\right)^2,\quad \al_+(\rho_1):=\left(1 + \rho_1^{-1/2}\right)^2.$$ 
In fact, $\al_-(\rho_1)$ and $\al_+(\rho_1)$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
\be\label{eq_deltaextra} 
 \delta(\hat v):=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)\hat v^2\lambda_{\min}^2(M)} \cdot  \norm{\Sigma_1^{1/2}(\beta_1 - \hat{v}\beta_2)}^2.\ee
%where $\lambda_{\min}(\hat M)$ is the smallest singular value of $\hat M$. 
Note that this factor converges to 0 as $\rho_1$ increases. It is not hard to see that Theorem \ref{thm_main_informal} is an immediate consequence of the following lemma.

@@

\subsection{Multiple Tasks}\label{app_proof_many_tasks}

@@

%In this section we consider the setting with $k$ many that have the same covariates.
%Since every task has the same number of data points as well as the same covariance, the only differences between different tasks are their models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
Using Lemma \ref{lem_minv} and some concentration bounds, we can complete the proof of Theorem \ref{thm_many_tasks}.

\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
In this setting, we need to study the following loss function:
\begin{align}
	f(B; W_1, \dots, W_t) = \sum_{i=1}^t \bignorm{X B W_i - Y_i}^2,\label{eq_mtl_same_cov}
\end{align}
for $W_1, W_2, \dots, W_t \in \R^r$. Then it is not hard to derive a closed form solution for $B$ as
	\begin{align*}
		\hat{B}(W_1, \dots, W_t) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^t Y_i W_i^{\top}} (\cal W  \cal W^{\top})^{-1} \\
		&= (B^\star \cal W ^{\top}) (\cal W \cal W ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^t \varepsilon_i W_i^{\top}} (\cal W \cal W^{\top})^{-1},
	\end{align*}
	where $\cal W \in\real^{r\times t}$ is a matrix defined as $\cal W=[W_1, W_2, \dots, W_t]$.
	%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
Then as in \eqref{approxvalid}, we pick $N$ independent samples of the training set for each task with $N\ge n^{1-\e_0}$, and use the concentration result, Lemma \ref{largedeviation}, to get the validation loss as 
\be\label{eq_multival}g(\cal W)=  N\left[\val(\cal W) + t  \sigma^2 \right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right).\ee

{\color{red}
\begin{align*}
	f(W_1, \dots, W_t) = \sum_{i=1}^t \bignorm{X \hat B W_i - Y_i}^2 = \val(\cal W) \cdot \left( 1+\OO(p^{-1/2+\e})\right),
\end{align*}
where
\begin{align*}
	 \val(\cal W) &= \sum_{i=1}^t  \bignorm{X\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2 + \E_{\e_i:1\le i \le t}\sum_{i=1}^t\left\| X(X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^t \varepsilon_i W_i^{\top}} (\cal W \cal W^{\top})^{-1}W_i -\e_i\right\| \\
	 &= \sum_{i=1}^t  \bignorm{X\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2 + \sigma^2 (nt-pr).
\end{align*}

}

Here $\val(\cal W)$ is defined as
%it remains to consider minimizing the validation loss
	$$\val( \cal W):=\exarg{\varepsilon_j, \forall 1\le j\le t}{ \sum_{i=1}^t \bignorm{\Sigma^{1/2}( \hat B W_i - \beta_i)}^2} =  \delta_{\bias}(\cal W) + \delta_{\vari} ,$$
where the model shift bias term $\delta_{\bias}(\cal W) $ is given by
	\begin{align*}
		\delta_{\bias}(\cal W) :=\sum_{i=1}^t  \bignorm{\Sigma^{1/2}\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2,
	\end{align*}
	and the variance term $\delta_{\vari} $ can be calculated as
	\begin{align*}
		\delta_{\vari} := \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
	\end{align*}
It suffices to minimize $\delta_{\bias}(\cal W)$ over $\cal W$, since both $t \sigma^2$ and $\delta_{\vari}$ do not depend on $\cal W$.

We denote $Q := \cal W^{\top} (\cal W\cal W^{\top})^{-1} \cal W \in\real^{t\times t}$, whose $(i,j)$-th entry is given by $W_i^{\top} (\cal W\cal W^{\top})^{-1} W_j$.
	%Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p \times k}$ denote the true model parameters.
	Now we can write $\delta_{\bias}(\cal W)$ succinctly as
	\begin{align*}
		\delta_{\bias}(Q) \equiv \delta_{\bias}(\cal W) = \bignormFro{\Sigma^{1/2}B^{\star}  \bigbrace{Q -\id}}^2 .
	\end{align*}
	From this equation we can solve the minimizer explicitly as $Q_0=U_{r}U_r^{\top}$. On the other hand, let $\hat {\cal W}$ be the minimizer of $g$, and denote $\hat Q:= \hat{\cal W}^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat{\cal W} $. We claim that $\hat Q$ satisfies
	\be\label{Q-Q}\|Q_0^{-1}\hat Q - \id\|_F = \oo(1) \quad \text{w.h.p.}\ee
	In fact, if \eqref{Q-Q} does not hold, then using the condition $\lambda_{\min}((B^{\star})^\top\Sigma B^{\star})\gtrsim \sigma^2$ and that $\delta_{\vari}(\cal W)=\OO(\sigma^2)$ by  Lemma \ref{lem_minv}, we obtain that
	$$   \val(\hat {Q}) + t  \sigma^2 > (\val( Q_0) + t \sigma^2 )\cdot (1+\oo(1)) \ \Rightarrow \ g( \hat{\cal W})>g( {\cal W}_0),$$
	where $\cal W_0\in \R^{r\times t}$ is a matrix such that  $ \cal W_0^{\top} (\cal W_0\cal W_0^{\top})^{-1} \cal W_0=Q_0$. Hence  $\hat {\cal W}$ is not a minimizer, which leads to a contradiction.

	In sum, we have solved that $\hat{\beta}_i^{\MTL}=B^{\star}\left( U_r v_i +\oo(1)\right)$. Inserting it into the definition of the test loss, we get that with high probability,
	\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) &= \bignorm{\Sigma^{1/2} \left((B^\star \hat{\cal W}^\top) (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t - \beta_t \right) }^2
		+ \sigma^2  \hat W_t^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t \cdot \bigtr{\Sigma (X^{\top}X)^{-1}} \\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \oo\left(\|B^\star\|^2\right) + \sigma^2\norm{v_t}^2 \bigtr{\Sigma (X^{\top}X)^{-1}} \cdot (1+\oo(1))\\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \frac{\sigma^2}{\rho-1}\norm{v_t}^2 + \oo \left( \|B^\star\|^2 + \sigma^2\right),
	\end{align*}
 where we used Lemma \ref{lem_minv} in the last step. On the other hand, by Lemma \ref{lem_minv} we have
	$$\te(\hat{\beta}_t^{\STL})=\frac{\sigma^2}{\rho-1} \cdot \left( 1+\oo(1)\right).$$
Combining the above two estimates, we conclude the proof.
\end{proof}
%From the above we can obtain three conceptual insights that are consistent with Section \ref{sec_denoise} and \ref{sec_insight}.
%\begin{itemize}
%	\item The de-noising effect of multi-task learning.
%	\item Multi-task training vs single-task training can be either positive or negative.
%	\item Transfer learning is better than the other two. And the improvement over multi-task training increases as the model distances become larger.
%\end{itemize}

>>>>

\section{Proof of the General Case}\label{sec_proof_general}

\subsection{Two-task Case}\label{app_proof_main_thm}

In this subsection, we give the proof Theorem \ref{thm_main_informal}. During the proof we shall also give explicit expressions for $\Delta_{\bias}$ and $\Delta_{\vari}$, which might be of interest to some readers.
With $a_i$, $i=1,2,3,4$, given in Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}, 
we define the following matrix
\be\label{defnpihat}\Pi \equiv \Pi(\hat v)= \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot \hat v^2{M} \frac{(1 +   a_3)\id + \hat v^2 a_4 {M}^{\top} {M}}{( \hat v^2 a_1 {M}^{\top} {M}+ a_2 )^2} {M}^{\top}.\ee
%which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
%$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
Moreover, we introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=\left(1- \rho_1^{-1/2}\right)^2,\quad \al_+(\rho_1):=\left(1 + \rho_1^{-1/2}\right)^2.$$ 
In fact, $\al_-(\rho_1)$ and $\al_+(\rho_1)$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
\be\label{eq_deltaextra} 
 \delta(\hat v):=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)\hat v^2\lambda_{\min}^2(M)} \cdot  \norm{\Sigma_1^{1/2}(\beta_1 - \hat{v}\beta_2)}^2.\ee
%where $\lambda_{\min}(\hat M)$ is the smallest singular value of $\hat M$. 
Note that this factor converges to 0 as $\rho_1$ increases. It is not hard to see that Theorem \ref{thm_main_informal} is an immediate consequence of the following lemma.

@@

\subsection{Proof of the Multiple-task Case}\label{app_proof_many_tasks}

@@

%In this section we consider the setting with $k$ many that have the same covariates.
%Since every task has the same number of data points as well as the same covariance, the only differences between different tasks are their models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
%Using Lemma \ref{lem_minv} and some concentration bounds, we can complete the proof of Theorem \ref{thm_many_tasks}.

\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
	We first introduce several notations.
	Let $\cal W = [W_1, W_2, \dots, W_t] \in \real^{r\times t}$ be a matrix that contains all the parameters of the output layers.
	Let $U_X U_X^{\top} = X (X^{\top} X)^{-1} X^{\top} \in\real^{n \times n}$ denote the projection matrix to $X$.

	Recall that all tasks have the same covariates in this setting.
	Therefore, the optimization objective equation \eqref{eq_mtl} becomes the following
	\begin{align}
		f(B; W_1, \dots, W_t) = \sum_{i=1}^t \bignorm{X B W_i - Y_i}^2, \label{eq_mtl_same_cov}
	\end{align}
	where $B \in \real^{p \times r}$ and $W_1, W_2, \dots, W_t \in \R^r$ for $1 < r < t$ by our assumption.
	Using the local optimality condition over $B$, that is $\nabla_B f = 0$, we obtain a closed form solution $\hat{B}$ that depends on the output layers as follows
	\begin{align*}
		\hat{B}(W_1, \dots, W_t) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^t Y_i W_i^{\top}} (\cal W  \cal W^{\top})^{-1} \\
		&= (B^\star \cal W ^{\top}) (\cal W \cal W ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top}   \bigbrace{\sum_{i=1}^t \varepsilon_i W_i^{\top}} (\cal W \cal W^{\top})^{-1}.
	\end{align*}
	Plugging in the equation $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} we obatin
	\begin{align}
		g(\cal W) = \sum_{j=1}^t \bignorm{X B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j + U_X U_X^{\top} \bigbrace{\sum_{i=1}^t \varepsilon_i W_i^{\top}} (\cW\cW)^{-1} W_j - Y_i}^2. \label{eq_mtl_output_layer}
	\end{align}
	Recall that we obtain $\hat{\cW}$ by minimizing $\cW$ in the above equation.
	In order to derive the minimizer, our proof involves two steps.
	First, we consider the expectation of $g(\cW)$ over $\varepsilon_1, \varepsilon_2, \dots, \varepsilon_t$, and $X$.
	We show that the minimizer of $\ex{g(\cW)}$ has a simple closed form solution similar to principal component analysis.
	Second, we show that the concentration error between $g(\cW)$ and $\ex{g(\cW)}$ is small provided with $n$ samples.
	Intuitively, this is because $\cW$ only has $r \times t$ parameters, which is much smaller than $n$ by our assumption.
	Hence we use standard concentration bounds and $\epsilon$-net arguments to show that the concentration error is small for $g(\cW)$.
	%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
	%Then as in \eqref{approxvalid}, we pick $N$ independent samples of the training set for each task with $N\ge n^{1-\e_0}$, and use the concentration result, Lemma \ref{largedeviation}, to get the validation loss as

	\medskip
	\noindent\textbf{The optimal solution of $\ex{g(\cW)}$.}
	We observe that $\ex{g(\cW)}$ admits a bias-variance decomposition as follows.
	First, we take expectation over $\varepsilon_1, \dots, \varepsilon_t$ conditional on $X$ and obtain the following:
	\begin{align*}
			\exarg{\varepsilon_1, \dots, \varepsilon_t}{g(\cW) \mid X}
		=& \bignorm{X B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} \cW - X B^{\star}}^2 \\
		&+ \sum_{j=1}^t\bigbrace{\exarg{\varepsilon_j}{\norm{\varepsilon_j}^2 - 2\inner{U_X U_X^{\top} {\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1} W_j}{\varepsilon_j}} } \\
		&+ \sum_{j=1}^t \exarg{\varepsilon_1, \dots, \varepsilon_t}{\bignorm{U_X U_X^{\top} \bigbrace{\sum_{i=1}^t  \varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1} W_j}^2} \\
		=& \bignorm{X B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} \cW - XB^{\star}}^2 + \sigma^2(n\cdot t - p\cdot r) 	\end{align*}
	The first equation uses the fact that $\varepsilon_i$ and $\varepsilon_j$ are pairwise independent for any $i \neq j$.
	The second equation uses the same fact and in addition uses that the variance of every entry of $\varepsilon_i$ is $\sigma^2$ for every $i = 1,\dots, t$.
	We further take expectation over $X$ and obtain the following:
	\begin{align*}
		\ex{g(\cW)} = n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 + \sigma^2 (n\cdot t - p \cdot r).
	\end{align*}
	%\be\label{eq_multival}g(\cal W)=  N\left[\val(\cal W) + t  \sigma^2 \right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right).\ee
\iffalse
{\color{red}
\begin{align*}
	f(W_1, \dots, W_t) = \sum_{i=1}^t \bignorm{X \hat B W_i - Y_i}^2 = \val(\cal W) \cdot \left( 1+\OO(p^{-1/2+\e})\right),
\end{align*}
where
\begin{align*}
	 \val(\cal W) &= \sum_{i=1}^t  \bignorm{X\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2 + \E_{\e_i:1\le i \le t}\sum_{i=1}^t\left\| X(X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^t \varepsilon_i W_i^{\top}} (\cal W \cal W^{\top})^{-1}W_i -\e_i\right\| \\
	 &= \sum_{i=1}^t  \bignorm{X\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2 + \sigma^2 (nt-pr).
\end{align*}

}
\fi
%Here $\val(\cal W)$ is defined as
%it remains to consider minimizing the validation loss
%	$$\val( \cal W):=\exarg{\varepsilon_j, \forall 1\le j\le t}{ \sum_{i=1}^t \bignorm{\Sigma^{1/2}( \hat B W_i - \beta_i)}^2} =  \delta_{\bias}(\cal W) + \delta_{\vari} ,$$
%where the model shift bias term $\delta_{\bias}(\cal W) $ is given by
%	\begin{align*}
%		\delta_{\bias}(\cal W) :=\sum_{i=1}^t  \bignorm{\Sigma^{1/2}\bigbrace{(B^\star \cal W^\top) (\cal W\cal W^{\top})^{-1} W_i - \beta_i}}^2,
%	\end{align*}
%	and the variance term $\delta_{\vari} $ can be calculated as
%	\begin{align*}
%		\delta_{\vari} := \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
%	\end{align*}
	One can see that the above equation provides a bias-variance decomposition of $\ex{g(\cW)}$.
	The first part measures the bias of multi-task learning compared to single-task learning for all tasks.
	The second part measures the variance of multi-task learning and scales with $\sigma^2$.
	In order to minimize $\ex{g(\cW)}$, it suffices to minimize the bias equation over $\cal W$, since the variance equation does not depend on $\cal W$.
	%We denote $Q := \cal W^{\top} (\cal W\cal W^{\top})^{-1} \cal W \in\real^{t\times t}$, whose $(i,j)$-th entry is given by $W_i^{\top} (\cal W\cal W^{\top})^{-1} W_j$.
	%Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p \times k}$ denote the true model parameters.
	%Now we can write $\delta_{\bias}(\cal W)$ succinctly as
	%\begin{align*}
	%	\delta_{\bias}(Q) \equiv \delta_{\bias}(\cal W) = \bignormFro{\Sigma^{1/2}B^{\star}  %\bigbrace{Q -\id}}^2 .
	%\end{align*}
	From the bias equation, we observe that the minimizer is equal to the best rank-$r$ (subspace) approximation to $B^{\star} \Sigma B^{\star}$, which is denoted by $U_{r} U_r^{\top}$.

	\medskip
	\noindent\textbf{The concentration error between $g(\cW)$ and $\ex{g(\cW)}$.}
	On the other hand, let $\hat {\cal W}$ be the minimizer of $g$, and denote $\hat Q:= \hat{\cal W}^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat{\cal W} $. We claim that $\hat Q$ satisfies
	\be\label{Q-Q}\|Q_0^{-1}\hat Q - \id\|_F = \oo(1) \quad \text{w.h.p.}\ee
	In fact, if \eqref{Q-Q} does not hold, then using the condition $\lambda_{\min}((B^{\star})^\top\Sigma B^{\star})\gtrsim \sigma^2$ and that $\delta_{\vari}(\cal W)=\OO(\sigma^2)$ by  Lemma \ref{lem_minv}, we obtain that
	$$   \val(\hat {Q}) + t  \sigma^2 > (\val( Q_0) + t \sigma^2 )\cdot (1+\oo(1)) \ \Rightarrow \ g( \hat{\cal W})>g( {\cal W}_0),$$
	where $\cal W_0\in \R^{r\times t}$ is a matrix such that  $ \cal W_0^{\top} (\cal W_0\cal W_0^{\top})^{-1} \cal W_0=Q_0$. Hence  $\hat {\cal W}$ is not a minimizer, which leads to a contradiction.

	\medskip
	\noindent\textbf{Putting both steps together.}
	In sum, we have solved that $\hat{\beta}_i^{\MTL}=B^{\star}\left( U_r v_i +\oo(1)\right)$. Inserting it into the definition of the test loss, we get that with high probability,
	\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) &= \bignorm{\Sigma^{1/2} \left((B^\star \hat{\cal W}^\top) (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t - \beta_t \right) }^2
		+ \sigma^2  \hat W_t^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t \cdot \bigtr{\Sigma (X^{\top}X)^{-1}} \\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \oo\left(\|B^\star\|^2\right) + \sigma^2\norm{v_t}^2 \bigtr{\Sigma (X^{\top}X)^{-1}} \cdot (1+\oo(1))\\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \frac{\sigma^2}{\rho-1}\norm{v_t}^2 + \oo \left( \|B^\star\|^2 + \sigma^2\right),
	\end{align*}
 where we used Lemma \ref{lem_minv} in the last step. On the other hand, by Lemma \ref{lem_minv} we have
	$$\te(\hat{\beta}_t^{\STL})=\frac{\sigma^2}{\rho-1} \cdot \left( 1+\oo(1)\right).$$
Combining the above two estimates, we conclude the proof.
\end{proof}
%From the above we can obtain three conceptual insights that are consistent with Section \ref{sec_denoise} and \ref{sec_insight}.
%\begin{itemize}
%	\item The de-noising effect of multi-task learning.
%	\item Multi-task training vs single-task training can be either positive or negative.
%	\item Transfer learning is better than the other two. And the improvement over multi-task training increases as the model distances become larger.
%\end{itemize}
